\renewcommand{\abstractname}{\Large Copyright}
\setlength{\parindent}{0pt}
\begin{abstract}
\setlength{\parindent}{0pt}
\nonzeroparskip

\begin{center}
\dwf, Version 4.1

Copyright\copyright 2010, 2013 \dwf\ Committee
\end{center}

\vspace{4ex}
Permission is granted to copy, 
distribute and/or modify this document under the terms of the
GNU Free Documentation License, Version 1.3; with no 
Invariant Sections, with no \mbox{Front-Cover}
Texts, and with no \mbox{Back-Cover} Texts. 

A copy of the license is included in the section entitled
``GNU Free Documentation License''.

This document is based in part on the \dwf, Version 2,
which contained the following notice:

\begin{myindentpara}{1cm}
UNIX International

Programming Languages SIG

Revision: 2.0.0 (July 27, 1993)

Copyright \copyright 1992, 1993 UNIX International, Inc.

Permission to use, copy, modify, and distribute this
documentation for any purpose and without fee is hereby
granted, provided that the above copyright notice appears
in all copies and that both that copyright notice and this
permission notice appear in supporting documentation, and
that the name UNIX International not be used in advertising or
publicity pertaining to distribution of the software without
specific, written prior permission. 
UNIX International makes
no representations about the suitability of this documentation
for any purpose. 
It is provided ``as is'' without express
or implied warranty.

\end{myindentpara}

This document is further based on the \dwf, Version 3, which
is subject to the GNU Free Documentation License.

Trademarks:

\begin{itemize}
\item Intel386 is a trademark of Intel Corporation.

\item Java is a trademark of Oracle, Inc.

\item All other trademarks found herein are property of their respective owners.

\end{itemize}

\end{abstract}
